---
title: "Statistical Analysis"
author: "Gert Vanhollebeke"
date: "06-01-2022"
output: html_document
---

```{css, echo=FALSE}
.codeblock {
  background-color: LightGray;
  border: 1px solid DimGray;
  font-weight: bold;
}
.outputblock {
  background-color: GhostWhite;
  border: 1px solid DimGray;
  font-weight: bold;
}
.commentstyle {
  font-weight: bold;
  color: black; 
}
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "F:/Study - EEGRUM/")
knitr::opts_chunk$set(class.source = "codeblock")
knitr::opts_chunk$set(class.output = "outputblock")
```

File which contains the statistical analysis for the article: TODO

**Data location:** TODO

**Code location:** TODO

# 1: Hypothesis 1

In this step, we research whether the total integration of a network can be explained by trait rumination scores of an individual.
In order to test this, we use linear regression where the questionnaire answers are the predictor variable and the total integration values are the response variable.


**Networks tested:**

- **DMN**: The default mode network (DMN), as described by Raichle (2015).  [(https://doi.org/10.1146/annurev-neuro-071013-014030)](https://doi.org/10.1146/annurev-neuro-071013-014030)
    
    - *Ventromedial Prefrontal Cortex* (Left/Right)
    - *Dorsomedial Prefrontal Cortex* (Left/Right)
    - *Posterior Cingulate Cortex* (Left/Right)
    - *Precuneus* (Left/Right)
    - *Angular Gyrus* (Left/Right)
    
<br>

- **Zhou**: The network found by Zhou and colleagues (2020) in their meta-analysis. [(https://doi.org/10.1016/j.neuroimage.2019.116287)](https://doi.org/10.1016/j.neuroimage.2019.116287)

    - *Anterior Cingulate Cortex* (Left)
    - *Precuneus* (Left)
    - *Superior Temporal Gyrus* (Left)
    - *Angular Gyrus* (Left)
    - *Inferior Frontal Gyrus* (Left)
    
<br>
    
- **Zhou Contralateral**: The contralateral version of the Zhou network.

    - *Anterior Cingulate Cortex* (Right)
    - *Precuneus* (Right)
    - *Superior Temporal Gyrus* (Right)
    - *Angular Gyrus* (Right)
    - *Inferior Frontal Gyrus* (Right)
    
<br>

**Frequency ranges tested: **


- **Alpha**: Alpha frequency range: 8 - 13 Hz

- **Beta**: Beta frequency range: 14 - 30 Hz

<br>

**All Combinations:**

- **DMN-Alpha**
- **DMN-Beta**
- **Zhou-Alpha**
- **Zhou-Beta**
- **ZhouCont-Alpha**
- **ZhouCont-Beta**

## 1.0: Load data

```{r data_loading}
#load questionnaire data
questionnaire_location <- 'Data/RStudio/Questionnaire_Data.csv'
questionnaire_data <- as.matrix(read.csv(questionnaire_location, TRUE))
RRS_data <- questionnaire_data[,c('RRS')] #keep only the RRS questionnaire data

#load the total integration data
TI_alpha_DMN <- as.matrix(read.csv('Data/RStudio/DMN_Alpha_Total_Integration.csv'))
TI_alpha_Zhou <- as.matrix(read.csv('Data/RStudio/Zhou_Alpha_Total_Integration.csv'))
TI_beta_DMN <- as.matrix(read.csv('Data/RStudio/DMN_Beta_Total_Integration.csv'))
TI_beta_Zhou <- as.matrix(read.csv('Data/RStudio/Zhou_Beta_Total_Integration.csv'))
TI_alpha_Zhou_Cont <- as.matrix(read.csv('Data/RStudio/Zhou_Contralateral_Alpha_Total_Integration.csv'))
TI_beta_Zhou_Cont <- as.matrix(read.csv('Data/RStudio/Zhou_Contralateral_Beta_Total_Integration.csv'))
```

## 1.1: Linear Regression

### 1.1.1: DMN-Alpha

```{r dmn_alpha}
dmn_alpha_predictor <- RRS_data
dmn_alpha_response <- TI_alpha_DMN
dmn_alpha_model <- lm(dmn_alpha_response ~ dmn_alpha_predictor)
summary(dmn_alpha_model)
```

### 1.1.2: DMN-Beta

```{r dmn_beta}
dmn_beta_predictor <- RRS_data
dmn_beta_response <- TI_beta_DMN
dmn_beta_model <- lm(dmn_beta_response ~ dmn_beta_predictor)
summary(dmn_beta_model)
```

### 1.1.3: Zhou-Alpha

```{r zhou_alpha}
zhou_alpha_predictor <- RRS_data
zhou_alpha_response <- TI_alpha_Zhou
zhou_alpha_model <- lm(zhou_alpha_response ~ zhou_alpha_predictor)
summary(zhou_alpha_model)
```

### 1.1.4: Zhou-Beta

```{r zhou_beta}
zhou_beta_predictor <- RRS_data
zhou_beta_response <- TI_beta_Zhou
zhou_beta_model <- lm(zhou_beta_response ~ zhou_beta_predictor)
summary(zhou_beta_model)
```

### 1.1.5: ZhouCont-Alpha

```{r zhoucont_alpha}
zhoucont_alpha_predictor <- RRS_data
zhoucont_alpha_response <- TI_alpha_Zhou_Cont
zhoucont_alpha_model <- lm(zhoucont_alpha_response ~ zhoucont_alpha_predictor)
summary(zhoucont_alpha_model)
```

### 1.1.6: ZhouCont-Beta

```{r zhoucont_beta}
zhoucont_beta_predictor <- RRS_data
zhoucont_beta_response <- TI_beta_Zhou_Cont
zhoucont_beta_model <- lm(zhoucont_beta_response ~ zhoucont_beta_predictor)
summary(zhoucont_beta_model)
```


## 1.2: Multiple testing correction and explanation

In order to conclude whether the RRS scores are predictive of the total integration of a specific network-frequency range combination. We look if any predictor coefficient has a corresponding significant p-value. The following p-values have been obtained:

- DMN-Alpha: 0.362
- DMN-Beta: 0.174
- Zhou-Alpha: 0.00477
- Zhou-Beta: 0.438
- ZhouCont-Alpha: 0.0793
- ZhouCont-Beta: 0.854

We apply false discovery rate (FDR) correction on these values to correct for multiple testing (Benjamini, Y., & Hochberg, Y., (1995))

```{r pvalue_correction_1}
#load pvalues
analysis_1_pvalues <- c(0.362, 0.174, 0.00477, 0.438, 0.0793, 0.854)
#correct pvalues
analysis_1_pvalues_corrected <- p.adjust(analysis_1_pvalues, "fdr")
#display pvalues
analysis_1_pvalues_corrected
```

We can see that after FDR correction the third analysis still remains significant. We can therefore conclude that the RRS values are predictive for the total integration in the network found by Zhou and colleagues (2020), but only in the alpha frequency range. Since the EEG data that is used is resting-state, eyes closed data, prominent alpha activity is normal and findings in this frequency range are to be expected.

### 1.3: Visualization and explanation of the findings

We will now look at the significant finding, the Zhou-alpha combination. We plot the summary of the model again:

```{r visualization}
summary(zhou_alpha_model)
```

We can see that the Zhou_alpha_predictor (the RRS values) is positive (= 0.028739). This means that a positive relationship is present, indicating that the higher the predictor value is (the RRS values), the higher the response value is (the Total Integration). This shows us that the total integration in the Zhou network increases when an individual scores higher on the RRS (rumination response scale). In laymans terms this means that if an individual ruminates more, the "Zhou network" has a higher total integration, which can be interpreted as more functionally connected. Finally we can see that the adjusted R-squared is 0.09551, indicating that roughly 10% of the variance in the data can be explained by the predictor.

Finally we plot the linear regression:

```{r regression plot}
#import library for visualization
library(ggplot2)
#build dataframe
regression.data <- data.frame(
                    predictor = zhou_alpha_predictor, 
                    response = zhou_alpha_response[,1]
                    )
#build plot of the predictor and response variable
regression_graph<-ggplot(regression.data, aes(predictor, response)) + #define data
                        geom_point() + #show points
                        geom_smooth(method="lm", col="black") + #plot regression line
                        ggtitle("Linear regression of the Zhou-Alpha network") + #plot title
                        theme(plot.title = element_text(hjust = 0.5)) + #center title
                        labs(x = "Predictor (RRS values)", y = "Response (Total Integration Values") #plot axis names

#plot graph
regression_graph

```

# 2: Hypothesis 2

Since we found a significant regression coefficient, we subsequently test whether the total integration of the network is driven by a single region in the network or not. To do this, we calculate the normalized degree of each region in the network. Normalized degree can be understood as calculating the degree of each region in the network (= sum of all connections involving this region), but normalizing this value by dividing it by the total integration of the network. This way we can see the contribution of each region to the total integration. The normalized degree can take any value between zero and 1.

Since the Zhou network has five regions, we will perform five tests. We now employ multivariate linear regression with the RRS values as predictor variable, and the normalized degree of each region as the response variables.

### 2.0: Load the data
```{r load data 2}
#RRS values are already loaded
#load normalized degree
Zhou_alpha_norm_degree <- as.matrix(read.csv('Data/RStudio/Normalized Degree (Alpha 8 - 13).csv'))
#Anterior Cingulate Cortex L norm. degree
norm_degree_ACC <- Zhou_alpha_norm_degree[,c("ACC_L")]
#Precuneus L norm. degree
norm_degree_Precun <- Zhou_alpha_norm_degree[,c("Precun_L")]
#Superior Temporal Gyrus L norm. degree
norm_degree_STG <- Zhou_alpha_norm_degree[,c("STG_L")]
#Angular Gyrus L norm. degree
norm_degree_AG <- Zhou_alpha_norm_degree[,c("AngGyr_L")]
#Inferior Frontal Gyrus L norm. degree
norm_degree_IFG <- Zhou_alpha_norm_degree[,c("IFG_L")]

#summarize the data
summary(Zhou_alpha_norm_degree)
```

### 2.1: Multivariate linear regression

```{r multivariate regression}
#define predictor
multivariate_predictor <- RRS_data
#build model
multivariate_model <- lm(cbind(norm_degree_ACC, norm_degree_Precun, norm_degree_STG, norm_degree_AG, norm_degree_IFG) ~ multivariate_predictor)
#summarize model
summary(multivariate_model)
```

### 2.2: Multiple testing correction and explanation

We have now five p-values, one for each normalized degree. These are the following:

- Anterior Cingulate Cortex L: 0.652
- Precuneus L: 0.0934
- Superior Temporal Gyrus L: 0.0121
- Angular Gyrus L: 0.116
- Inferior Frontal Gyrus L: 0.0664

We apply false discovery rate (FDR) correction on these values to correct for multiple testing (Benjamini, Y., & Hochberg, Y., (1995))

```{r multiple correction 2}
#load pvalues
analysis_2_pvalues <- c(0.652, 0.0934, 0.0121, 0.116, 0.0664)
#correct pvalues
analysis_2_pvalues_corrected <- p.adjust(analysis_2_pvalues, "fdr")
#display pvalues
analysis_2_pvalues_corrected
```

We see that while the initial p-value for the coefficient of the Superior Temporal Gyrus was significant (p = 0.0121), it becomes insignificant after multiple testing correction. We can therefore conclude that the total integration is not driven by the change in normalized degree of an individual region. Replication of these findings should be conducted to solidify these initial findings. The close-to-significant result of the Superior Temporal Gyrus might also be further explained in this replication.